---
title: "data_reform"
author: "Yuting Qiu"
date: "3/13/2020"
output: html_document
---

---
title: "recode 0 as misisng"
output: html_notebook
author: "Yuting Qiu"
date: "2020-03-12"
---

Description: the goal of this file is to recode the 0 in the original file as missing data since and converge the files again since we dotn't know if the 0s are really 0 or not inculated

```{r}
library(readxl)
# both TBBC3 population and Z025 population are in it
C19_measure <- read_excel("CFAR200_2019-Plant.xlsx")
M19_measure <- read_excel("MF-700_2019_Plant.xlsx")
# there are two locations for one population
TBBC3_19_plot <- read_excel("19DN_plots.xlsx")
Z025_19_plot <- read_excel("19DR_plots.xlsx")

tbbc3_18_mp <- read_excel("18DN_plot_measurements.xlsx")
```


```{r}
# seperate the two populations measurement from one table
C19_Z025_measure <- C19_measure[C19_measure$`Experiment Name` == "19DR_rating",]
C19_TBBC3_measure <- C19_measure[C19_measure$`Experiment Name` == "19DN_rating",]
M19_Z025_measure <- M19_measure[M19_measure$`Experiment Name` == "19DR_rating",]
M19_TBBC3_measure <- M19_measure[M19_measure$`Experiment Name` == "19DN_rating",]

# seperate the two locations from the plot table
TBBC3_C19_plot <- TBBC3_19_plot[TBBC3_19_plot$Field == "CFAR200",]
TBBC3_M19_plot <- TBBC3_19_plot[TBBC3_19_plot$Field == "MF-700",]
Z025_C19_plot <- Z025_19_plot[Z025_19_plot$Field == "CFAR200",]
Z025_M19_plot <- Z025_19_plot[Z025_19_plot$Field == "MF-700",]
```


```{r}
# join the plot data with the measurement data
# more data in the meansurement file, use side join
# need to create a new column first for measurement file
C19_TBBC3_measure$Plot <- substr(C19_TBBC3_measure$`Plant ID`,1,8)
C19_Z025_measure$Plot <- substr(C19_Z025_measure$`Plant ID`,1,8)
M19_TBBC3_measure$Plot <- substr(M19_TBBC3_measure$`Plant ID`,1,8)
M19_Z025_measure$Plot <- substr(M19_Z025_measure$`Plant ID`,1,8)

z025_c19_mp <- merge(C19_Z025_measure,Z025_C19_plot, by.x = "Plot", by.y = "Plot ID", all.x=TRUE)
tbbc3_m19_mp <- merge(M19_TBBC3_measure,TBBC3_M19_plot,by.x = "Plot", by.y = "Plot ID", all.x = TRUE)
z025_m19_mp <- merge(M19_Z025_measure,Z025_M19_plot,by.x = "Plot", by.y = "Plot ID", all.x = TRUE)
tbbc3_c19_mp <- merge(C19_TBBC3_measure,TBBC3_C19_plot,by.x = "Plot", by.y = "Plot ID", all.x = TRUE)
```

```{r}
# clean the dataframe
z025_c19_mp1 <- z025_c19_mp[,c(1,3,5,6,7,9,13,22,23)]
z025_m19_mp1 <- z025_m19_mp[,c(1,3,5,6,7,9,13,22,23)]
tbbc3_c19_mp1 <- tbbc3_c19_mp[,c(1,3,5,6,7,9,13,22,23)]
tbbc3_m19_mp1 <- tbbc3_m19_mp[,c(1,3,5,6,7,9,13,22,23)]
```

# the real missing data need to be dealed for 2018
```{r}
tbbc3_18_mp[tbbc3_18_mp$value == ".",3] <- NA # replace the . with NA
tbbc3_18_mp1 <- tbbc3_18_mp[,c(1,3,6,9,13,14,25,26)]
tbbc3_18_mp1 <- na.omit(tbbc3_18_mp1)
```


# look at the 2018 dataset
```{r}
# seperate the count from the rating
tbbc3_18_mp1_c <- tbbc3_18_mp1[tbbc3_18_mp1$measurement_parameter_id =="StandCount",]
tbbc3_18_mp1_r <- tbbc3_18_mp1[tbbc3_18_mp1$measurement_parameter_id =="DiseasePCTRating1",]

# see if any line has little number of count that cannot be used
summary(tbbc3_18_mp1_c)
tbbc3_18_mp1_c$value <- as.numeric(tbbc3_18_mp1_c$value) # count from 0 to 21
tbbc3_18_mp1_c <- tbbc3_18_mp1_c[order(tbbc3_18_mp1_c$value),]
tbbc3_18_less <- head(tbbc3_18_mp1_c,12)
vc <- tbbc3_18_less$plot_id

# get the ratings of those lines
tbbc3_18_less_r <- tbbc3_18_mp1_r[tbbc3_18_mp1_r$plot_id %in% vc,]
# delete those information
tbbc3_18_mp1_r <- tbbc3_18_mp1_r[!tbbc3_18_mp1_r$plot_id %in% vc,]
```


# 2019 data reformatting 
# remove the real missing data
```{r}
# seperate the coverage from the severity
tbbc3_c19_mp1$coverage <- substr(tbbc3_c19_mp1$value,1,1)
tbbc3_c19_mp1$severity <- substr(tbbc3_c19_mp1$value,2,2)
tbbc3_m19_mp1$coverage <- substr(tbbc3_m19_mp1$value,1,1)
tbbc3_m19_mp1$severity <- substr(tbbc3_m19_mp1$value,2,2)
z025_c19_mp1$coverage <- substr(z025_c19_mp1$value,1,1)
z025_c19_mp1$severity <- substr(z025_c19_mp1$value,2,2)
z025_m19_mp1$coverage <- substr(z025_m19_mp1$value,1,1)
z025_m19_mp1$severity <- substr(z025_m19_mp1$value,2,2)

tbbc3_c19_mp1$coverage <- as.numeric(tbbc3_c19_mp1$coverage)
tbbc3_m19_mp1$coverage <- as.numeric(tbbc3_m19_mp1$coverage)
z025_c19_mp1$coverage <- as.numeric(z025_c19_mp1$coverage)
z025_m19_mp1$coverage <- as.numeric(z025_m19_mp1$coverage)

# remove the real missing data in the dataframe
tbbc3_c19_mp1 <- na.omit(tbbc3_c19_mp1)
tbbc3_m19_mp1 <- na.omit(tbbc3_m19_mp1)
z025_c19_mp1 <- na.omit(z025_c19_mp1)
z025_m19_mp1 <- na.omit(z025_m19_mp1)
```

# convert the 0-100% scale of 2018 data to 0-8 scale and to 1-9 scale with missing for 2018
```{r}
library(dplyr)
tbbc3_18_conv <- tbbc3_18_mp1_r
tbbc3_18_conv$value <- as.numeric(as.character(tbbc3_18_conv$value))
tbbc3_18_conv$value.cvt <- NA
tbbc3_18_conv$value.cvt1 <- NA

# need to write a loop
for (i in 1:nrow(tbbc3_18_conv)) {
   if (tbbc3_18_conv[i,2] == 100) {
    tbbc3_18_conv[i,9] = 8
  } else if (tbbc3_18_conv[i,2] > 89 & tbbc3_18_conv[i,2]<100){
    tbbc3_18_conv[i,9] = 7
  } else if (tbbc3_18_conv[i,2] > 75 & tbbc3_18_conv[i,2]<90){
    tbbc3_18_conv[i,9] = 6
  } else if (tbbc3_18_conv[i,2] > 60 & tbbc3_18_conv[i,2]<76) {
    tbbc3_18_conv[i,9] = 5
  } else if (tbbc3_18_conv[i,2] > 45 & tbbc3_18_conv[i,2]<61) {
    tbbc3_18_conv[i,9] = 4
  } else if (tbbc3_18_conv[i,2] > 30 & tbbc3_18_conv[i,2]<46) {
    tbbc3_18_conv[i,9] = 3
  } else if (tbbc3_18_conv[i,2] > 15 & tbbc3_18_conv[i,2]<31) {
    tbbc3_18_conv[i,9] = 2
  } else if (tbbc3_18_conv[i,2] > 0 & tbbc3_18_conv[i,2]<16) {
    tbbc3_18_conv[i,9] = 1
  } else if (tbbc3_18_conv[i,2] == 0){
    tbbc3_18_conv[i,9] = 0
  } 
}


# convert the 0-100% scale to 1-8 scale with 0 as missing 
for (i in 1:nrow(tbbc3_18_conv)) {
   if (tbbc3_18_conv[i,2] == 100) {
    tbbc3_18_conv[i,10] = 8
  } else if (tbbc3_18_conv[i,2] > 89 & tbbc3_18_conv[i,2]<100){
    tbbc3_18_conv[i,10] = 7
  } else if (tbbc3_18_conv[i,2] > 75 & tbbc3_18_conv[i,2]<90){
    tbbc3_18_conv[i,10] = 6
  } else if (tbbc3_18_conv[i,2] > 60 & tbbc3_18_conv[i,2]<76) {
    tbbc3_18_conv[i,10] = 5
  } else if (tbbc3_18_conv[i,2] > 45 & tbbc3_18_conv[i,2]<61) {
    tbbc3_18_conv[i,10] = 4
  } else if (tbbc3_18_conv[i,2] > 30 & tbbc3_18_conv[i,2]<46) {
    tbbc3_18_conv[i,10] = 3
  } else if (tbbc3_18_conv[i,2] > 15 & tbbc3_18_conv[i,2]<31) {
    tbbc3_18_conv[i,10] = 2
  } else if (tbbc3_18_conv[i,2] > 0 & tbbc3_18_conv[i,2]<16) {
    tbbc3_18_conv[i,10] = 1
  } else if (tbbc3_18_conv[i,2] == 0){
    tbbc3_18_conv[i,10] = NA
  } 
}
```

# convert the 0 as missing for 2019 dataset
```{r}
tbbc3_c19_mp1$coverage_1 <- tbbc3_c19_mp1$coverage
tbbc3_m19_mp1$coverage_1 <- tbbc3_m19_mp1$coverage
z025_c19_mp1$coverage_1 <- z025_c19_mp1$coverage
z025_m19_mp1$coverage_1 <- z025_m19_mp1$coverage


tbbc3_c19_mp1$coverage_1[tbbc3_c19_mp1$coverage_1 ==0] <- NA
tbbc3_m19_mp1$coverage_1[tbbc3_m19_mp1$coverage_1 ==0] <- NA
z025_c19_mp1$coverage_1[z025_c19_mp1$coverage_1 == 0] <- NA
z025_m19_mp1$coverage_1[z025_m19_mp1$coverage_1 == 0] <- NA

# it is okay to combine the two locations together for a popualation
tbbc3_19 <- rbind(tbbc3_c19_mp1, tbbc3_m19_mp1)
z025_19 <- rbind(z025_c19_mp1, z025_m19_mp1)
```


# i can check the distribution of check line now
```{r}
checks <- c("B73","M162W","M37W","NC344","Oh7B","Tx303")

tbbc3_c19_check <- tbbc3_c19_mp1[tbbc3_c19_mp1$`Plot Name` %in% checks,]
boxplot(coverage_1 ~ tbbc3_c19_check$`Plot Name`, data =tbbc3_c19_check, main = "Boxplot of check lines in 2019 TBBC3 population in CFAR200", xlab = "check lines", ylab = "disease coverage")

tbbc3_m19_check <- tbbc3_m19_mp1[tbbc3_m19_mp1$`Plot Name` %in% checks,]
boxplot(coverage_1 ~ tbbc3_m19_check$`Plot Name`, data =tbbc3_m19_check, main = "Boxplot of check lines in 2019 TBBC3 population in MF700", xlab = "check lines", ylab = "disease coverage")

z025_m19_check <- z025_m19_mp1[z025_m19_mp1$`Plot Name` %in% checks,]
boxplot(coverage_1 ~ z025_m19_check$`Plot Name`, data=z025_m19_check, main = "Boxplot of check lines in 2019 Z025 population in MF700", xlab = "check lines", ylab = "disease coverage")

z025_c19_check <- z025_c19_mp1[z025_c19_mp1$`Plot Name` %in% checks,]
boxplot(coverage_1 ~ z025_c19_check$`Plot Name`, data=z025_c19_check, main = "Boxplot of check lines in 2019 Z025 population in CFAR200", xlab = "check lines", ylab = "disease coverage")

tbbc3_18_check <- tbbc3_18_conv[tbbc3_18_conv$plot %in% checks,]
boxplot(value.cvt1 ~ tbbc3_18_check$plot, data=tbbc3_18_check, main = "Boxplot of check lines in 2018 TBBC3 population", xlab = "check lines", ylab = "disease coverage")
```
# the plots are not really the case, if i use the mean to plot, things can change, beacuse there can have real 0 that for the whole plot being coeded as na



# calcualte the mean, incidence rate mummified rate of each plot in 2018 dataset
## convert to 1-8 scale and count 
### cvt -- converted to 1-8 scale
```{r}
library(dplyr)
# get the mean of tbbc3 population in 2018 
tbbc3_18_mean <- aggregate(value.cvt~plot_id, data =tbbc3_18_conv, mean)
tbbc3_18_mean1 <- aggregate(value.cvt1~plot_id, data =tbbc3_18_conv, mean)
tbbc3_18_mean %>% rename(mean = value.cvt) -> tbbc3_18_mean # 108 obs
tbbc3_18_mean1 %>% rename(mean1 = value.cvt1) -> tbbc3_18_mean1 # 105 obs due to the missing data
tbbc3_18_mean <- merge(x=tbbc3_18_mean,y=tbbc3_18_mean1, all.x = TRUE) # 108 obs in this way
# the 0 in the mean 1 should be the true 0
tbbc3_18_mean$mean1[is.na(tbbc3_18_mean$mean1)] <- 0

# get the number of count of each plot (genotype) ONLY ONE REP and for each sepecfic value in the plot
tbbc3_18_conv %>% 
  group_by(plot_id) %>% 
  mutate(plot_count=n()) -> tbbc3_18_conv

# do not count the na as plot number
tbbc3_18_conv %>% 
  group_by(plot_id) %>% 
  summarise(non_na_count = sum(!is.na(value.cvt1))) -> tbbc3_18_pc # plot count without the na
tbbc3_18_conv <- merge(x=tbbc3_18_conv, y=tbbc3_18_pc, by.x = "plot_id", by.y = "plot_id", all.x = TRUE)

# count for each of the rating
tbbc3_18_conv %>% 
             group_by(plot_id,value.cvt) %>% 
             mutate(value.cvt_cout = n()) -> tbbc3_18_conv

# get the percentage (rate) of each count in r
tbbc3_18_conv %>% 
  mutate(value.cvt_rate = value.cvt_cout/plot_count) -> tbbc3_18_conv
tbbc3_18_conv %>% 
  mutate(value.cvt_rate1 = value.cvt_cout/non_na_count) -> tbbc3_18_conv

# find the incidence rate and the mummified rate
# incidence date = 1 - value rate with 0%
tbbc3_18_conv$value <- as.numeric(tbbc3_18_conv$value)
tbbc3_18_conv$value.cvt <- as.numeric(tbbc3_18_conv$value.cvt)

tbbc3_18_incidence <- tbbc3_18_conv[tbbc3_18_conv$value.cvt==0,]
tbbc3_18_incidence <- unique(tbbc3_18_incidence)
tbbc3_18_incidence$incidence_rate <- 1 - tbbc3_18_incidence$value.cvt_rate
# if we regard the 0 as na, we wont have incidence rata data out of it
# the inf in the data should be the 0 actually, because the non_na_count is 0, means everything in that plot 0, which can be regareded as the true 0, besides, the plot with rating less than 5 had already beening deleted.

# mummified rate # use 100%, or 8 as the indicator of totally mummified
tbbc3_18_mummified <- tbbc3_18_conv[tbbc3_18_conv$value.cvt == 8,]
tbbc3_18_mummified <- na.omit(tbbc3_18_mummified)
tbbc3_18_mummified <- unique(tbbc3_18_mummified)
rename(tbbc3_18_mummified, mummi_rate = value.cvt_rate) -> tbbc3_18_mummified
# if we really need mummified 1? yes, we do, the data is missing and we have no means knowing if that is 0 or 8
rename(tbbc3_18_mummified, mummi_rate1 = value.cvt_rate1) -> tbbc3_18_mummified
# many missing data here due to the all 0. 

# combine the three together
# in the mummified, if there is NA, it should be 0 
# in the indicence, if there is NA, it should be 100% (we cannot find any value being 0)
# combined mean with incidence
tbbc3_18_agg <- merge(x=tbbc3_18_mean,y=tbbc3_18_incidence, by.x = "plot_id", by.y = "plot_id", all.x = TRUE, all.y = TRUE)
tbbc3_18_agg$incidence_rate[is.na(tbbc3_18_agg$incidence_rate)] <- 1

# aggregate with mummified
tbbc3_18_agg <- merge(x=tbbc3_18_agg,y=tbbc3_18_mummified, by.x = "plot_id", by.y = "plot_id", all.x = TRUE, all.y = TRUE)
tbbc3_18_agg$mummi_rate[is.na(tbbc3_18_agg$mummi_rate)] <- 0
tbbc3_18_agg$mummi_rate1[is.na(tbbc3_18_agg$mummi_rate1)] <- 0

# merge the information of different tables
tbbc3_18_agg$value <- coalesce(tbbc3_18_agg$value.x, tbbc3_18_agg$value.y)
tbbc3_18_agg$value.cvt <- coalesce(tbbc3_18_agg$value.cvt.x, tbbc3_18_agg$value.cvt.y)
tbbc3_18_agg$row_num <- coalesce(tbbc3_18_agg$row_num.x, tbbc3_18_agg$row_num.y)
tbbc3_18_agg$plot <- coalesce(tbbc3_18_agg$plot.x, tbbc3_18_agg$plot.y)
tbbc3_18_agg$rep <- coalesce(tbbc3_18_agg$rep.x, tbbc3_18_agg$rep.y)
tbbc3_18_agg$range_num <- coalesce(tbbc3_18_agg$range_num.x, tbbc3_18_agg$range_num.y)
tbbc3_18_agg$block <- coalesce(tbbc3_18_agg$block.x, tbbc3_18_agg$block.y)
tbbc3_18_agg$plot_count <- coalesce(tbbc3_18_agg$plot_count.x, tbbc3_18_agg$plot_count.y)


tbbc3_18_agg1 <- tbbc3_18_agg[,c(1,2,3,18,31:40)]
tbbc3_18_agg1 <- unique(tbbc3_18_agg1)
# if the mean of a plot is NA, that plot is all 0, it is true 0 in a way

cor.test(tbbc3_18_agg1$mean, tbbc3_18_agg1$mummi_rate, method = "pearson") #0.9256085 
cor.test(tbbc3_18_agg1$mean, tbbc3_18_agg1$incidence_rate, method = "pearson") # 0.8953216
cor.test(tbbc3_18_agg1$mummi_rate, tbbc3_18_agg1$incidence_rate, method = "pearson") # 0.7619616

cor.test(tbbc3_18_agg1$mean1, tbbc3_18_agg1$mummi_rate1, method = "pearson") #0.8753698 

# our correlation coefficient is significantly lower in this analysis: in last time, I put all check lines together, but this time, the check lines are seperated may I need to remove the checks to see. 
# remove check lines
tbbc3_18_agg2 <- tbbc3_18_agg1[!tbbc3_18_agg$plot %in% c("B73","NC344","Tx303"),]
cor.test(tbbc3_18_agg2$mean, tbbc3_18_agg2$mummi_rate, method = "pearson") # 0.925564
cor.test(tbbc3_18_agg2$mean, tbbc3_18_agg2$incidence_rate, method = "pearson") # 0.8591019 
cor.test(tbbc3_18_agg2$mummi_rate, tbbc3_18_agg2$incidence_rate, method = "pearson") #0.7285691

cor.test(tbbc3_18_agg2$mean1, tbbc3_18_agg2$mummi_rate1, method = "pearson") # 0.8857184
```


# 2019 data Z025 population MF
```{r}
# count the datapoint we have for each point. 
# get the number of count of each plot_id for Z025 population in 2019 in MF (2 replication in total)

z025_19 %>% 
  group_by(Plot) %>% 
  mutate(plot_count=n()) -> z025_19

z025_19 %>% 
  group_by(Plot) %>% 
  summarise(non_na_count = sum(!is.na(coverage_1))) -> z025_19_pc # plot count without the na
z025_19 <- merge(x=z025_19, y=z025_19_pc, by.x = "Plot", by.y = "Plot", all.x = TRUE)

# a count of 5 would be used as a cutoff 
z025_19_1 <- z025_19[z025_19$plot_count >= 5,]

# calculate the mean for each plot
z025_19_1 %>%
  group_by(Plot) %>%
  mutate(mean = mean(coverage)) -> z025_19_1

z025_19_1 %>%
  group_by(Plot) %>%
  mutate(mean1 = mean(coverage_1,na.rm=TRUE)) -> z025_19_1


# count the number of each value in the rate
z025_19_1 %>%
  group_by(Plot, coverage) %>%
  mutate(coverage_count = n()) -> z025_19_1

z025_19_1 %>%
  group_by(Plot, coverage_1) %>%
  mutate(coverage_count1 = n()) -> z025_19_1

# get the percentage of each count in a plot
z025_19_1 %>% 
  mutate(coverage_rate = coverage_count/plot_count) -> z025_19_1
# for non-na count 
z025_19_1 %>% 
  mutate(coverage_rate1 = coverage_count1/non_na_count) -> z025_19_1

# find the incidence rate and mummified rate for each plot
z025_19_incidence <- z025_19_1[z025_19_1$coverage == 0,]
z025_19_incidence <- unique(z025_19_incidence[,-2])
z025_19_incidence$incidence_rate <- 1-z025_19_incidence$coverage_rate
# for the 0 coded as na, the incidence rate would be 1 

z025_19_mummi <- z025_19_1[z025_19_1$coverage == 8,]
z025_19_mummi <- unique(z025_19_mummi[,-2])
rename(z025_19_mummi, mummi_rate=coverage_rate) -> z025_19_mummi
rename(z025_19_mummi, mummi_rate1=coverage_rate1) -> z025_19_mummi #0 code as na

# merge the three table together
# combine the three together
# in the mummified, if there is NA, it should be 0 
# in the indicence, if there is NA, it should be 100% (we cannot find any value being 0)

# combined mean with mummified
z025_19_agg <- merge(x=z025_19_1,y=z025_19_mummi[,c(1,18,19)], by.x = "Plot", by.y = "Plot", all.x = TRUE, all.y = TRUE)
z025_19_agg$mummi_rate[is.na(z025_19_agg$mummi_rate)] <- 0
z025_19_agg$mummi_rate1[is.na(z025_19_agg$mummi_rate1)] <- 0
# agrregate with incidence
z025_19_agg <-  merge(x=z025_19_agg,y=z025_19_incidence[,c(1,20)], by.x = "Plot", by.y = "Plot", all.x = TRUE, all.y = TRUE)
z025_19_agg$incidence_rate[is.na(z025_19_agg$incidence_rate)] <- 1

```


#convert the severity in z025 populaiton and get mean for each plot
```{r}
# replace the severity with three levels "L(less) U(usual) M(more)"
z025_19_agg$severity_new <- NA
for (i in 1:nrow(z025_19_agg)) {
  if (z025_19_agg[i,11] == "+"){
    z025_19_agg[i,24] = "3"
  } else if (z025_19_agg[i,11] == "-"){
    z025_19_agg[i,24] = "1"
  } 
  else {z025_19_agg[i,24] = "2"}
}
z025_19_agg$severity_new <- as.numeric(as.character(z025_19_agg$severity_new))

z025_19_agg %>%
  group_by(Plot) %>%
  mutate(severity_mean = mean(severity_new)) -> z025_19_agg

z025_19_agg1 <- z025_19_agg
z025_19_agg1$mean1[is.nan(z025_19_agg1$mean1)] <- 0
# remove the check lines and calculate the correlation coefficient

z025_19_agg2 <- z025_19_agg1[!z025_19_agg$`Plot Name` %in% c("M37W","M162W","Tx303","B73"),]
cor.test(z025_19_agg2$mean, z025_19_agg2$mummi_rate, method = "pearson") # 0.3259189 
cor.test(z025_19_agg2$mean, z025_19_agg2$incidence_rate, method = "pearson") # 0.8673955 
cor.test(z025_19_agg2$mummi_rate, z025_19_agg2$incidence_rate, method = "pearson") # 0.198947  
# all significant, but signidicant is not important

cor.test(z025_19_agg2$mean1, z025_19_agg2$mummi_rate1, method = "pearson") #0.314348  

cor.test(z025_19_agg2$mean, z025_19_agg2$severity_mean, method = "pearson") # 0.3738682  
cor.test(z025_19_agg2$incidence_rate, z025_19_agg2$severity_mean, method = "pearson") # 0.1040342 
# there is some correlation between the coverage and severity for eacg plot
```


# for tbbc3 population
```{r}
# remove the plot with less than 5 readings
tbbc3_19 %>% 
  group_by(Plot) %>% 
  mutate(plot_count=n()) -> tbbc3_19

tbbc3_19 %>% 
  group_by(Plot) %>% 
  summarise(non_na_count = sum(!is.na(coverage_1))) -> tbbc3_19_pc # plot count without the na
tbbc3_19 <- merge(x=tbbc3_19, y=tbbc3_19_pc, by.x = "Plot", by.y = "Plot", all.x = TRUE)

# a count of 5 would be used as a cutoff 
tbbc3_19_1<- tbbc3_19[tbbc3_19$plot_count >= 5,]

# calculate the mean for each plot
tbbc3_19_1 %>%
  group_by(Plot) %>%
  mutate(mean = mean(coverage)) -> tbbc3_19_1
# the mean for 0 as na
tbbc3_19_1 %>%
  group_by(Plot) %>%
  mutate(mean1 = mean(coverage_1,na.rm=TRUE)) -> tbbc3_19_1

# count the number of each value in the rate
tbbc3_19_1 %>%
  group_by(Plot, coverage) %>%
  mutate(coverage_count = n()) -> tbbc3_19_1

tbbc3_19_1 %>%
  group_by(Plot, coverage_1) %>%
  mutate(coverage_count1 = n()) -> tbbc3_19_1

# get the percentage of each count in a plot
tbbc3_19_1 %>% 
  mutate(coverage_rate = coverage_count/plot_count) -> tbbc3_19_1
tbbc3_19_1 %>% 
  mutate(coverage_rate1 = coverage_count1/non_na_count) -> tbbc3_19_1

# find the incidence rate and mummified rate for each plot
tbbc3_19_incidence <- tbbc3_19_1[tbbc3_19_1$coverage == 0,]
tbbc3_19_incidence <- unique(tbbc3_19_incidence[,-2]) #col 2 indicate the plant in a plot
tbbc3_19_incidence$incidence_rate <- 1-tbbc3_19_incidence$coverage_rate

tbbc3_19_mummi <- tbbc3_19_1[tbbc3_19_1$coverage == 8,]
tbbc3_19_mummi <- unique(tbbc3_19_mummi[,-2]) # very few 8 ratings
rename(tbbc3_19_mummi, mummi_rate=coverage_rate) -> tbbc3_19_mummi
rename(tbbc3_19_mummi, mummi_rate1=coverage_rate1) -> tbbc3_19_mummi

# merge the three table together
# combine the three together
# in the mummified, if there is NA, it should be 0 
# in the indicence, if there is NA, it should be 100% (we cannot find any value being 0)

# combined mean with incidence
tbbc3_19_agg <- merge(x=tbbc3_19_1,y=tbbc3_19_incidence[,c(1,20)], by.x = "Plot", by.y = "Plot", all.x = TRUE, all.y = TRUE)
tbbc3_19_agg$incidence_rate[is.na(tbbc3_19_agg$incidence_rate)] <- 1

# aggregate with mummified
tbbc3_19_agg <- merge(x=tbbc3_19_agg,y=tbbc3_19_mummi[,c(1,18,19)], by.x = "Plot", by.y = "Plot", all.x = TRUE, all.y = TRUE)
tbbc3_19_agg$mummi_rate[is.na(tbbc3_19_agg$mummi_rate)] <- 0
tbbc3_19_agg$mummi_rate1[is.na(tbbc3_19_agg$mummi_rate1)] <- 0
```


# recode the severity and calculate the mean 
```{r}
tbbc3_19_agg$severity_new <- NA

# replace the severity with three levels "L(less) U(usual) M(more)"
for (i in 1:nrow(tbbc3_19_agg)) {
  if (tbbc3_19_agg[i,11] == "+"){
    tbbc3_19_agg[i,24] = "3"
  } else if (tbbc3_19_agg[i,11] == "-"){
    tbbc3_19_agg[i,24] = "1"
  } 
  else {tbbc3_19_agg[i,24] = "2"}
}

tbbc3_19_agg$severity_new <- as.numeric(as.character(tbbc3_19_agg$severity_new))

tbbc3_19_agg %>%
  group_by(Plot) %>%
  mutate(severity_mean = mean(severity_new)) -> tbbc3_19_agg

# if the mean is 0, then the mean1 should also be 0, it is the real 0
tbbc3_19_agg1 <- tbbc3_19_agg
tbbc3_19_agg1$mean1[is.nan(tbbc3_19_agg1$mean1)] <- 0


# remove the check lines and calculate the correlation coefficient
tbbc3_19_agg2 <- tbbc3_19_agg1[!tbbc3_19_agg$`Plot Name` %in% checks,]
cor.test(tbbc3_19_agg2$mean, tbbc3_19_agg2$mummi_rate, method = "pearson") # 0.09024929 
cor.test(tbbc3_19_agg2$mean, tbbc3_19_agg2$incidence_rate, method = "pearson") # 0.8282704  
cor.test(tbbc3_19_agg2$mummi_rate, tbbc3_19_agg2$incidence_rate, method = "pearson") #0.04993093   
# all significant, but signidicant is not important

cor.test(tbbc3_19_agg2$mean1, tbbc3_19_agg2$mummi_rate1, method = "pearson") # 0.1243563   

cor.test(tbbc3_19_agg2$mean, tbbc3_19_agg2$severity_mean, method = "pearson") # 0.2322353   
# there is some correlation between the coverage and severity for eacg plot
```



# write out the results
```{r}
write.csv(tbbc3_18_agg1, "/Users/qiuyuting/Documents/data_file/tbbc3_18_agg.csv")
write.csv(z025_19_agg1, "/Users/qiuyuting/Documents/data_file/z025_19_agg.csv")
write.csv(tbbc3_19_agg1, "/Users/qiuyuting/Documents/data_file/tbbc3_19_agg.csv")
```

